---
title: "Distributions homework"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

You are given a cut of a database containing details of mobile phones and their customer reviews on an online shopping site in file `20190928-items.csv`. Each row in the data corresponds to one model of mobile phone. Use the data to answer the following questions.  

As part of the MVP we want you not to just run the code but also have a go at **intepreting the results** - this is the **main learning objective** from this homework. Don't worry if you aren't sure, it's ok to get it wrong, but give it a go, through trying we will learn.  

# MVP

## Question 1

Load the data, `clean_names()` and explore it.

```{r}
library(janitor)
library(tidyverse)

phones <- read_csv("data/20190928-items.csv") %>% clean_names()
```


<br>

## Question 2

Find the brand with the highest numbers of phones in the dataset.

```{r}
phones %>% 
  group_by(brand) %>% 
  summarise(count = n()) %>% 
  slice_max(count)
```


<br>

## Question 3

For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

<br>
[**Hint** - you will need to calculate the `mean()` and `sd()` of the ratings to plot the appropriate normal. Alternatively, investigate using the `fitdistr()` function in the `MASS` package]

```{r}
phones <- phones %>% 
  select(brand, rating) %>% 
  filter(brand == "Samsung") 
```

```{r}
# "VISUALIZE your data before you do any statistics!"
phones %>% 
  ggplot() +
  aes(x = rating) +
  geom_histogram(col = "white", bins = 25)
```

```{r}
# Is the sample distribution normal?

## q-q plot
library(ggpubr)

ggqqplot(phones$rating)

## OR explicitly Shapiro-Wilks' method
## H0: sample distribution is normal;
## H1: sample distribution is not normal
## p-value = 1.088e-12
## p-value <<<, test is not significant, the sample distribution is normal
shapiro.test(phones$rating) 
```


```{r}
phones_stats <- phones %>% 
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating)
  )

phones_stats
```

```{r}
phones %>% 
  ggplot() +
  aes(x = rating) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
  stat_function(
    fun = dnorm,
    args = list(
      mean = phones_stats$mean,
      sd = phones_stats$sd
    ),
  col = "red"
  )
```


<br>

## Question 4

We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the `mean()` and `sd()` of your distribution in an earlier question. Use these values to calculate the *proportions* of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.  

```{r}
# the values are normally distributed, I can use standardized variable (z-score) to evaluate how far away we are from the mean
## creating new column displaying z-score
## creating new column with three categories depending on the z-score
phones_z_score_SD <- phones %>% 
  mutate(
    z_score_rating = scale(rating)
  ) %>% 
  mutate(
    how_far_from_the_mean = case_when(
      z_score_rating <= -3 | z_score_rating >= 3 ~ "3 SD",
      z_score_rating <= -2 | z_score_rating >= 2 ~ "2 SD",
      TRUE                                       ~ "1 SD",
    )
  )

phones_z_score_SD

# *proportions* of ratings within one-, two- and three standard deviations away of the mean 
## displayed in frequency table
freq_table_phones <- phones_z_score_SD$how_far_from_the_mean %>% 
  tabyl() %>% 
  as_tibble()

freq_table_phones

## displayed in a graph
freq_table_phones %>% 
  ggplot() +
  aes(x = ., y = percent) +
  geom_col() +
  theme_light() +
  labs(
    x = "none",
    title = "Proportion of phone ratings",
    subtitle = "within one-, two- and three standard deviations away of the mean"
  )
```



# Extension

Investigate the use of the `qqnorm()` plot in R to determine whether data is normally distributed. What is a 'quantile-quantile plot'? What do we expect to see in a `qqnorm()` plot if data is nearly normally distributed?

Here are a few resources:

* [Understanding QQ plots](https://data.library.virginia.edu/understanding-q-q-plots/)
* [Short video tutorial](https://www.youtube.com/watch?v=Erze9pNIX8A) (the presenter doesn't use RStudio, but the commands are the same. Just ignore the final comments on hypothesis tests for now.)
* [Longer video tutorial](https://www.youtube.com/watch?v=smJBsZ4YQZw)
* [Quantile-Quantile plots - R Base Graphs](http://www.sthda.com/english/wiki/qq-plots-quantile-quantile-plots-r-base-graphs)

