---
title: "R Notebook"
output: html_notebook
---

1 MVP
The file project_management.csv contains data sampled from the recent work schedule of a small construction company. Column estimated_length contains the estimated length of a building job in days, while column actual_length contains the actual recorded length of the job in days.

We are interested in determining the accuracy of the job estimations made by the company using simple linear regression, so we will eventually want to run a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable.



Load the data into a dataframe project

```{r}
library(tidyverse)
library(janitor)
library(broom)
library(ggfortify)

data <- read_csv("data/project_management.csv")
```


Plot the data, taking estimated_length as the independent variable and actual_length as the dependent variable.

```{r}
data %>% 
  ggplot() +
  aes(x = estimated_length, y = actual_length) +
  geom_point()
```

Calculate the correlation coefficient of estimated_length and actual_length and interpret the value you obtain.

```{r}
data %>% 
  summarise(correlation = cor(estimated_length, actual_length))
```

Perform a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable. Save the model object to a variable.

```{r}
model <- lm(formula = actual_length ~ estimated_length, data = data)

model
clean_names(glance(model))
```

Interpret the regression coefficient of estimated_length (i.e. slope, gradient) you obtain from the model. How do you interpret the r2
 value reported by the model?

```{r}
# actual_length = b0 + b1 * estimated_length

# b0 = 1.416 fitted model will intercept y axis at 1.416
# b1 = 1.223 slope (gradient) of the model line. With every 1 day increase in estimated_length, the actual_length will increase by 1.223 days.
# r^2 = 0.6474772	(coeff. of determination) explains how good fit our model is. Not so good.. low amount of explained variation. Approximately 65% of the variation in actual_length can be predicted from the variation in estimated_length
```

Is the relationship statistically significant? Remember, to assess this you need to check the p
-value of the regression coefficient (or slope/gradient). But you should first check the regression diagnostic plots to see if the p
-value will be reliable (don’t worry about any outlier points you see in the diagnostic plots, we’ll return to them in the extension).

```{r}
autoplot(model)
# p-value = 1.911284e-12	< 0.005 (0.001), we reject H0; r^2 = 0
# the relationship is statistically significant
```

