---
title: "Advanced `ggplot2` Homework - Answers"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

# Learning Objectives<br>

* Practice making plots with ggplot2
* Create your own theme and use it with plots

# Background

QikBit is a start-up company producing low cost fitness trackers. Their flagship product, the KwikBit Unit1 (spot the mistake on the patent form) reproduces the functionality of the best-in-class commercial fitness trackers, but is cheap to make.  They are currently preparing to pitch to a potential investor and require a selection of charts to include in their presentation and promotional materials.

# Data

The data for this assignment is provided in the `CodeClanData` package.

* `qb_revenue_breakdown`
* `qb_monthly_sales`
* `qb_competitors`
* `qb_device_data`

Take a little time before you start to familiarise yourself with the data.

```{r}
library(tidyverse)
library(CodeClanData)
library(scales)
```

```{r}
head(qb_revenue_breakdown) 
head(qb_monthly_sales)
head(qb_competitors)
head(qb_device_data)
```


# MVP

## Part 1

1. Take the data in the dataset `qb_revenue_breakdown` and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(qb_revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_bar(position = "stack", stat = "identity") +
  theme_light() +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(
    values = c(
      "Accessories" = "#E89FE9",
      "Smart Watches" = "#50434F",
      "Subscription Services" = "#B6A7B5",
      "Trackers" = "#F9A472"
    )
  ) +
  labs(
    title = "Source of revenue in 2018 and 2019"
  )
```


2.	Make a line chart showing monthly sales of the "flagship product" - the 'KwikBit Unit 1' - over the last year. This data is inside `qb_monthly_sales`, but you will need to use `filter()` or `subset()` (base `R)`.

```{r}
qb_monthly_sales %>% 
  filter(str_detect(as.character(Date), "2019")) %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  separate(
    col = Date,
    into = c("year", "month"),
    sep = "-0"
  ) %>% 
  ggplot() +
    aes(x = month, y = Amount, group = Cashflow, colour = Cashflow) +
    geom_line(colour = "#50434F") +
    theme_light() +
    labs(
      y = "amount ($)",
      title = "Monhly sales of KwikBit Unit 1 in 2019"
    )
```


3.	Make a line chart showing monthly revenue and costs over the last year. This data is also in `qb_monthly_sales`.				

```{r}
qb_monthly_sales %>% 
  filter(str_detect(as.character(Date), "2019")) %>% 
  filter(Cashflow == "Total Revenue" | Cashflow == "Cost") %>% 
  separate(
    col = Date,
    into = c("year", "month"),
    sep = "-0"
  ) %>% 
  ggplot() +
    aes(x = month, y = Amount, group = Cashflow, colour = Cashflow) +
    geom_line() +
    theme_light() +
    scale_y_continuous(labels = scales::comma) +
    scale_colour_manual(
      values = c(
      "Cost" = "#E89FE9",
      "Total Revenue" = "#50434F")
    ) +
    labs(
      y = "amount ($)",
      title = "Monthly revenue and costs over the last year"
    )
```


4.	Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use `geom_area`). This data is in `qb_competitors`.

```{r}
ggplot(qb_competitors) +
  aes(x = Year, y = Revenue, group = Company, fill = Company) +
  geom_area() +
  scale_y_continuous(labels = scales::comma) +
  theme_light() +
  scale_fill_manual(
    values = c(
      "Competitor 1" = "#E89FE9",
      "Competitor 2" = "#50434F",
      "Competitor 3" = "#B6A7B5",
      "Competitor 4" = "#F9A472"
    )
  ) +
  labs(
    y = "Revenue ($)",
    title = "Annual sales of personal fitness trackers"
  )
```


5. Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
ggplot(qb_competitors) +
  aes(x = Year, y = Revenue, group = Company, colour = Company) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  theme_light() +
  scale_colour_manual(
    values = c(
      "Competitor 1" = "#E89FE9",
      "Competitor 2" = "#50434F",
      "Competitor 3" = "#B6A7B5",
      "Competitor 4" = "#F9A472"
    )
  ) +
  labs(
    y = "Revenue ($)",
    title = "Annual sales of personal fitness trackers"
  )
```


6. Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in `qb_device_data`.

```{r}
qb_device_data_hours <- qb_device_data %>% 
  group_by(device, id, hours) %>% 
  mutate(
    number_of_steps = sum(counts)
  ) %>% 
  summarise(number_of_steps = sum(counts))


ggplot(qb_device_data_hours) +
  aes(x = hours, y = number_of_steps, group = id, col = id) +
  geom_line() +
  facet_wrap(~ device, labeller = "label_both") +
  theme_light() +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(
    values = c(
      "person1" = "#E89FE9",
      "person2" = "#50434F",
      "person3" = "#B6A7B5",
      "person4" = "#F9A472",
      "person5" = "#BD7040"
    )
  ) +
  labs(
    x = "time (hours)",
    y = "number of steps",
    title = "Steps measured by different devices"
  )

# First I created a new column 'number_of_steps' that summarises the number of steps in each hour, grouped by device and person.
# Then I plotted a line graph showing the number of steps over time and faceted it into four plots by devices.
# We are getting similar step counts across all four devices.
```


## Part 2 

Take the plots that you produced in part one and now polish them by:

1. Adding appropriate labels
2. Changing the scales and coordinates when appropriate.
3. Applying a unified theme, which is described below:
  * Graphs should have white backgrounds, and use colour sparingly.
  * There should be faint grid lines.
  * Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller. 
  * All plots should use colours from the following company colour scheme.

```{r}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```


# Extension

Try out different ways of plotting these three datasets. Which graphs do you think are particularly effective at showing QikBit's success?
