---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

1 MVP

### Load the diamonds.csv data set and undertake an initial exploration of the data. You will find a description of the meanings of the variables on the relevant [Kaggle page](https://www.kaggle.com/shivam2503/diamonds/)

```{r}
library(tidyverse)
library(mosaic)
library(mosaicData)
library(ggfortify)
library(GGally)
library(fastDummies)

diamonds <- read_csv("data/diamonds.csv")
```

### We expect the carat of the diamonds to be strong correlated with the physical dimensions x, y and z. Use ggpairs() to investigate correlations between these four variables.

```{r}
diamonds %>% 
  ggpairs(columns = c("carat", "x", "y", "z")
  )

# there is a very strong correlation between the carat and the physical dimensions (x, y and z)
```

### So, we do find significant correlations. Let’s drop columns x, y and z from the dataset, in preparation to use only carat going forward.

```{r}
diamonds <- diamonds %>% 
  select(-c(x, y, z))
```

### We are interested in developing a regression model for the price of a diamond in terms of the possible predictor variables in the dataset.

### Use ggpairs() to investigate correlations between price and the predictors (this may take a while to run, don’t worry, make coffee or something).

```{r}
ggpairs(diamonds)
```

### Perform further ggplot visualisations of any significant correlations you find.

```{r}
# there is also very strong correlation between the carat and price
diamonds %>% 
  summarise(correlation_coef = cor(carat, price))

diamonds %>% 
  ggplot() +
  aes(x = carat, y = price) +
  geom_point() 
```

### Shortly we may try a regression fit using one or more of the categorical predictors cut, clarity and color, so let’s investigate these predictors:
### Investigate the factor levels of these predictors. How many dummy variables do you expect for each of them?

```{r}
unique(diamonds$cut) # 4 dummy variables
unique(diamonds$clarity) # 7 
unique(diamonds$color) # 6

# expected dummy variables: n-1
```

### Use the dummy_cols() function in the fastDummies package to generate dummies for these predictors and check the number of dummies in each case.

```{r}
diamonds %>% 
  dummy_cols(select_columns = c("cut", "clarity", "color"),
             remove_first_dummy = TRUE,
             remove_selected_columns = TRUE)
```




