---
title: "R Notebook"
output: html_notebook
---

1 MVP

### You are given a set of data on housing sale prices for the last few years in King County (near Seattle) between May 2014 and May 2015.


### We want you to build an **explanatory model for the price of housing** in King County, i.e. an interpretable model in which the included variables are statistically justifiable.

The variable definitions are:

id - Unique ID for each home sold
date - Date of the home sale
price - Price of each home sold
bedrooms - Number of bedrooms
bathrooms - Number of bathrooms, where .5 accounts for a room with a toilet but no shower
sqft_living - Square footage of the apartments interior living space
sqft_lot - Square footage of the land space
floors - Number of floors
waterfront - A dummy variable for whether the apartment was overlooking the waterfront or not
view - An index from 0 to 4 of how good the view of the property was
condition - An index from 1 to 5 on the condition of the apartment
grade - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design
sqft_above - The square footage of the interior housing space that is above ground level
sqft_basement - The square footage of the interior housing space that is below ground level
yr_built - The year the house was initially built
yr_renovated - The year of the house’s last renovation
zipcode - What zipcode area the house is in
lat - Lattitude
long - Longitude
sqft_living15 - The square footage of interior housing living space for the nearest 15 neighbors
sqft_lot15 - The square footage of the land lots of the nearest 15 neighbors

```{r}
library(tidyverse)
library(fastDummies)
library(GGally)
```

```{r}
houses <- read_csv("data/kc_house_data.csv")

glimpse(houses)
```
### Tidy up the data ready for regression:

### You might like to think about removing some or all of date, id, sqft_living15, sqft_lot15 and zipcode (lat and long provide a better measure of location in any event).
### Have a think about how to treat waterfront. Should we convert its type? 
  #### I think it already behaves like a dummy, and it is numerical, no need to transform it.. But I have to transform it in order to be able to run ggpairs(), split the data set into two (numerical and logical) so I don't run out of memory
### We converted yr_renovated into a renovated logical variable, indicating whether the property had ever been renovated. You may wish to do the same.
### Have a think about how to treat condition and grade? Are they interval or categorical ordinal data types? 
  #### both condition and grade are categorical ordinal variables, I should dummy them to transform them into numerical variables, but frankly, lm() does it for me..
  #### For simplification, I transformed grade into three supersets (3 ordinal categories - low/average/high)

```{r}
houses_tidy <- houses %>% 
  select(-c("id", "date", "zipcode", "sqft_living15", "sqft_lot15")) %>% 
  mutate(
    waterfront = as.logical(waterfront)
  ) %>% 
  mutate(
    renovated = ifelse(yr_renovated == "0", 0, 1) %>%  as.logical()
  ) %>% 
  select(-yr_renovated) %>%
  mutate(
    condition = factor(condition, levels = c("1", "2", "3", "4", "5")),
    grade = case_when(
      grade %in% c(1:3) ~ "low",
      grade %in% c(4:10) ~ "average",
      TRUE ~ "high"
    ),
    grade = factor(grade, levels = c("low", "average", "high"))
  )
  # dummy_cols(
  #   select_columns = c("waterfront", "condition", "grade"),
  #   remove_first_dummy = TRUE,
  #   remove_selected_columns = TRUE
  # ) 

glimpse(houses_tidy)
```

### Check for aliased variables using the alias() function (this takes in a formula object and a data set). [Hint - formula price ~ . says ‘price varying with all predictors’, this is a suitable input to alias()]. Remove variables that lead to an alias. Check the ‘Elements of multiple regression’ lesson for a dropdown containing further information on finding aliased variables in a dataset.

```{r}
# check for redundant variables
alias(lm(price ~ ., data = houses_tidy))
```

```{r}
# get rid of identified redundant variable
houses_tidy <- houses_tidy %>% 
  select(-sqft_basement)
```

### Systematically build a regression model containing up to four main effects (remember, a main effect is just a single predictor with coefficient), testing the regression diagnostics as you go

### splitting datasets into numeric and non-numeric columns might help ggpairs() run in manageable time, although you will need to add either a price or resid column to the non-numeric dataframe in order to see its correlations with the non-numeric predictors.

```{r}
houses_tidy_numeric <- houses_tidy %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- houses_tidy %>%
  select_if(function(x) !is.numeric(x))
```

```{r}
houses_tidy_numeric %>% 
  ggpairs(progress = FALSE) # no progress bar
```

```{r}
houses_tidy_nonnumeric$price <- houses_tidy$price

houses_tidy_nonnumeric %>% 
  ggpairs(progress = FALSE)
```

The highest correlation between numeric explanatory variables and `price` has variable `sqft_living`

## Model 1

```{r}

```

