---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Homework Info

This homework recaps what you learnt in class today. 

You'll be looking at the `backpack` dataset contained within the `CodeClanData` package using tidyverse. You'll need to load the appropriate packages first.     

**In all plots you make, ensure your labels are appropriate. You also must write down a short summary of what each plot tells you in order to meet MVP.**  

**Make sure you create a new project in which to do this homework!**


# MVP

<br>

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions.   

```{r}
library(tidyverse)
library(CodeClanData)

head(backpack)
glimpse(backpack)
names(backpack)
dim(backpack)

unique(backpack$back_problems) # just 0 and 1, will need to change to factor

backpack <- backpack %>% 
  mutate(
    back_problems = factor(back_problems,
                           levels = c("0", "1"),
                           labels = c("Yes", "No"))
  )
```



**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 


```{r}
ggplot(backpack) +
  aes(x = backpack_weight) +
  geom_histogram(fill = "steel blue", color = "dark blue") +
  scale_x_continuous(breaks = seq(0, 40, by = 5)) +
  scale_y_continuous(breaks = seq(0, 13, by = 2)) +
  labs(
    x = "Backpack weight",
    y = "Number of backpacks",
    title = "Distribution of backpack weights carried by students"
  )

# This is a histogram showing the distribution of backpack weight carried by students.
# The horizontal x axis represents the Backpack weight
# The vertical y axis shows the count of backpacks of each weight category
# I changed the scale of the both axes to a bit more finely-grainded scale
```


**Question 3.**  

Are there differences in backpack weight between sex?   

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, fill = sex) +
  geom_histogram(col = "white") +
  scale_x_continuous(breaks = seq(0, 40, by = 10)) +
  scale_y_continuous(breaks = seq(0, 13, by = 2)) +
  facet_wrap(~ sex) +
  labs(
    x = "Backpack weight",
    y = "Number of backpacks",
    title = "Distribution of backpack weights carried by students",
    subtitle = "Depending on the students' gender"
  )

backpack %>% 
  group_by(sex) %>% 
  summarise(avg_weight = mean(backpack_weight, rm.na = TRUE))

# There is just a slight difference in the backpack weights between the genders. 
# Male carry in average 0.6 more then females
```



**Question 4.**  

What is the relationship between body weight and backpack weight?

```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight) +
  geom_point() +
  geom_smooth(method='lm', se = FALSE) +
  labs(
    x = "Body weight",
    y = "Backpack weight",
    title = "Relationship between body weight and backpack weight"
  )

backpack %>% 
  lm(formula = body_weight ~ backpack_weight) %>% 
  summary.lm

# This is a scatter plot showing the relationship between body weight and backpack weight.
## body weight is an independent variable here, while backpack weight is dependent (response) variable.
# A second layer of a smoothed straight trend line geom, aka linear regression, was applied. This showed a positive correlation between the two variables, the regression slope is positive. (in general, the higher the body weight is, the higher the backpack weight is).

# Using the linear regression formula (equation Y = a + b * X), we got the intersect point and the slope respectively as: backpack weight = 141.79 + 0.97 * body weight.
## the positive slope number shows the positive correlation in exact numbers
```



**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  

```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight) +
  geom_point(
    aes(colour = sex)) +
  geom_smooth(method='lm', se = FALSE) +
  facet_wrap(~ sex) +
  labs(
    x = "Body weight",
    y = "Backpack weight",
    title = "Relationship between body weight and backpack weight",
    subtitle = "Depending on the students' gender"
  )

backpack %>% 
  filter(sex == "Female") %>% 
  lm(formula = body_weight ~ backpack_weight) %>% 
  summary.lm

backpack %>% 
  filter(sex == "Male") %>% 
  lm(formula = body_weight ~ backpack_weight) %>% 
  summary.lm

# These two scatter plots represent the relationship between body weight and backpack weight depending on gender.
# Both plots have got a liner regression line drawn into them to show the "steepness" of the line
# The facet_wrap function respects the scale. Thus it is clear that female students tend to carry heavier backpacks the higher their body weight is compared to male students.
# Both are showing a positive correlation tho, female just steeper (slope 0.84 vs 0.41)
```


**Question 6.**  

What **proportion** of students report back pain?   

```{r}
ggplot(backpack) +
  aes(x = back_problems) +
  geom_bar(fill = c("salmon", "turquoise")) +
  labs(
    x = "Back problem",
    y = "Number of students",
    title = "Proportion of students with and without a back pain"
  )

backpack %>% 
  group_by(back_problems) %>% 
  summarise(count = n())

# This is a simple bar chart representing the proportion of students with and without a back pain.
# There is more than twice as many students with reported back pain than those without.
## note. The colors were set to match the colors scheme for gender in R
```



**Question 7.**  

Is this proportion different between men and women? 

```{r}
ggplot(backpack) +
  aes(x = back_problems, fill = sex) +
  geom_bar() +
  facet_wrap(~ sex) +
  labs(
    x = "Back problem",
    y = "Number of students",
    title = "Proportion of students with and without a back pain",
    subtitle = "Depending on the students' gender"
  )

# This is two bar charts showing the proportion of students with and without a back pain depending on their gender.
# Clearly, males suffer from back pain more often than females.
```



**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.   

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, y = body_weight, fill = sex) +
  geom_bar(stat = "identity") +
  facet_grid(sex ~ back_problems, labeller = "label_both") +
  labs(
    x = "Back weight",
    y = "Body weight",
    title = "Relationship between backpack weight, body weight, gender and reported back pain"
  )

# This plot describes the relationship between backpack weight, body weight, gender and reported back pain.
# The plot consists of four bar charts. 
## The left ones focuses on the students experiencing back pain, whilst the right ones on students with no reported back pain.
## The top ones represent female students, the bottom ones the male students. 
# The variables of continuous type (back weight, body weight) were mapped to the aesthetic attributes of the plot.
```


# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data.  

```{r}
backpack_no_zero <- backpack %>% 
  filter(year != 0) %>% 
  arrange(year)

ggplot(backpack_no_zero) +
  aes(x = year, fill = back_problems) +
  geom_bar() +
  scale_x_continuous(breaks = 1:6) +
  scale_y_continuous(breaks = seq(0, 30, by = 5)) +
  labs(
    x = "Year",
    y = "Number of students",
    title = "Back pain across the year groups"
  )

# Students from the year group 0 were dropped.
# From this bar chart there isn't any significant difference in reported back pain across the year groups
```



**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?

```{r}
backpack %>% 
  group_by(year) %>% 
  summarise(avg_backpack_weight = mean(backpack_weight, na.rm = TRUE)) %>% 
  ggplot() +
  aes(x = year, y = avg_backpack_weight) +
  geom_point() +
  scale_x_continuous(breaks = 1:6) +
  #scale_y_continuous(breaks = seq(0, 30, by = 5)) +
  labs(
    x = "Year",
    y = "Backpack weight",
    title = "How heavy the bags those in different years carry"
  )

# First I calculated the average weight of a backpack in each year and the plotted it using the scatter graph, where the independent x variable represents group year and the dependent y variable represents the average backpack weight.
# Average weight across the group years looks rather random.
```


