---
title: "R Notebook"
output: html_notebook
---

Clustering homework
You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings.

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

Perform k-means clustering and chose a value of k.

Visualise the clustering for your chosen value of k.

Do you think the clustering seems a good fit for this data?

Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).

```{r}
library(tidyverse)
library(janitor)
library(corrplot)
library(broom)
library(factoextra)
library(cluster)

mall_customers <- read_csv("data/mall_customers.csv") %>% clean_names()
```

### 1. Do some exploratory analysis on the data. Comment on findings.

```{r}
mall_customers %>%
  select(gender) %>% 
  group_by(gender) %>% 
  summarise(n = n())
```

```{r}
# plot correlation matrix
mall_customers_numeric <- mall_customers %>% 
  select(-c(customer_id, gender))
  
corrplot(cor(mall_customers_numeric), method = "number", type = "lower")
```

```{r}
mall_customers %>% 
  ggplot() +
  aes(x = age, y = annual_income_k) +
  geom_point()

mall_customers %>% 
  ggplot() +
  aes(x = age, y = spending_score_1_100) +
  geom_point()
```

Findings: 
* we have 200 observations in the sample, there is more females than males (122/88). I removed the `gender` column as k means clustering only work on numeric data
* variables `age` and `annual_income_k` do not seem to have any relation between them, `age` and `spending_score_1_100` seem to have a slight negative correlation, and finally, no correlation between the `annual_income_k` and `spending_score_1_100` neither

### 2. We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

```{r}
# tidy the data, k-means clustering method can only work on numeric data
mall_customers <- mall_customers %>% 
  column_to_rownames("customer_id") %>% 
  select(-c("gender", "age"))

head(mall_customers)
```

```{r}
# standardize data
mall_customers_scale <- mall_customers %>% 
  mutate_if(is.numeric, scale)

# check it worked - OK
mall_customers_scale %>%
  as_tibble() %>%
  pivot_longer(cols = c(annual_income_k, spending_score_1_100), 
               names_to = "type", 
               values_to = "value") %>% #convert data to long format
  group_by(type)%>%
  summarise(mean = round(mean(value)), 
            sd = sd(value))
```

### 3. Perform k-means clustering and chose a value of k.

#### 1. Elbow method

```{r}
# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(mall_customers_scale, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, mall_customers_scale)
  )

k_clusters
```
```{r}
clusterings <- k_clusters %>%
  unnest(glanced)

ggplot(clusterings, aes(x=k, y=tot.withinss)) +
  geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1))
```

k = 5 levels the grade off the most, maybe?

### 2. Silhouette coefficient

```{r}
#Chosen k=10 as an example 
cluster_list_k10 <-  clusterings %>% 
  unnest(augmented) %>%
  filter(k == 10) %>%
   select(.cluster) %>%
    pull()
  
sil <- silhouette(as.numeric(cluster_list_k10), 
                  dist(mall_customers_scale))

fviz_silhouette(sil)
```

All of the clusters have a score above the average (red dotted line), difficult to tell what k would be the best from this graph, it is quite ambiguous 

```{r}
fviz_nbclust(mall_customers_scale, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)
```

k = 5 suggested here again

#### 3. Gap statistic

```{r}
fviz_nbclust(mall_customers_scale, 
             kmeans, 
             method = "gap_stat", 
             nstart = 25, 
             k.max = 10)
```

it is pointing toward k = 5 again, we will go with that

### 4. Visualise the clustering for your chosen value of k.
k = 5

```{r}
clusterings %>% 
  unnest(augmented)
```


```{r}
clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 5) %>%
  ggplot(aes(x = annual_income_k, y = spending_score_1_100, colour = .cluster)) +
    geom_point(aes(color = .cluster)) 
```

### 5. Do you think the clustering seems a good fit for this data?

yes, looking good! The 5 clusters is well distinguished

### 6. Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).
