---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Market basket analysis (MBA) uses
association rules mining looking at items bought together in a single "basket" purchase
The association rules used for MBA use the probability principles 

Three values
- support
- confidence
- lift

```{r}
library(tidyverse)
library(arules)
library(arulesViz)

transactions <- read_csv("data/online_retail_subset.csv")
```

```{r}
head(transactions)
```

# calculating "on paper"

For the first section we are interested in the purchase of two particular items:

item A - ‘HEART OF WICKER SMALL’ (StockCode 22469)
item B - ‘LARGE CAKE TOWEL PINK SPOTS’ (StockCode 21110)

1. Calculate the support for item A
 (this will be the support for a single item)
 
```{r}
transactions %>% 
  filter(StockCode == 22469) %>% 
  nrow()

nrow(transactions)

# sup(A)
110/30000
```

2. Calculate the support and confidence for rule (A→B)

```{r}
# keeping only those transactions that contained both A and B in "one basket"
# 5
transactions %>% 
  select(InvoiceNo, StockCode, Description) %>% 
  filter(StockCode == 22469 | StockCode == 21110) %>% 
  arrange(InvoiceNo) %>% 
  mutate(InvoiceNo_lag = lag(InvoiceNo)) %>% 
  mutate(same_invoice = ifelse(InvoiceNo == InvoiceNo_lag, "yes", "no")) %>% 
  mutate(Description_lag = lag(Description)) %>% 
  mutate(same_item = ifelse(Description == Description_lag, "yes", "no")) %>% 
  filter(same_invoice == "yes" & same_item == "no") %>% 
  nrow()
  
# sup(A→B)
5/30000

# conf(A→B)
0.00017/0.0037
```

3. Calculate the lift for (A→B)

```{r}
transactions %>% 
  filter(StockCode == 21110) %>% 
  nrow()

nrow(transactions)

# sup(B)
14/30000

# lift(A→B)
0.00017/(0.0037 * 0.00047)
```


# calculating using R package arules

Values:

sup(A) 0.0037
sup(B) 0.00047
sup(A→B)  0.00017
conf(A→B) 0.046
lift(A→B) 98

```{r}
transactions_reformat <- transactions %>%
  select(InvoiceNo, Description) %>%
  drop_na()

write_csv(transactions_reformat, "data/transactions_reformat.csv")

apriori_format <- read.transactions("data/transactions_reformat.csv", format = "single", sep = ",", header = TRUE, cols = c("InvoiceNo", "Description"))

class(transactions)
```

```{r}
head(apriori_format)
inspect(head(apriori_format))
```


```{r}
# Create an item frequency plot for the top 20 items
itemFrequencyPlot(apriori_format,topN=20,type="absolute")
```


```{r}
# pass the minimum required support and confidence
# Get the rules
# set the left hand side to be “HEART OF WICKER SMALL” and find its antecedents
rules <- apriori(apriori_format, parameter = list(supp = 0.0001, conf = 0.2),
                 appearance = list(default = "rhs", lhs = "HEART OF WICKER SMALL"), # right hand side/left hand side
                 control = list(verbose = FALSE))

rules <- sort(rules, decreasing = TRUE, by = "confidence")

inspect(rules)

summary(rules)

# interpretation: if someone buys HEART OF WICKER SMALL, they are 42% likely to buy HEART OF WICKER LARGE too
```

```{r}
plot(rules,  
     shading = "confidence")
```





